---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# qualR

<!-- badges: start -->
<!-- badges: end -->

The goal of qualR is facilitate the download of air pollutants and meteorological information from 
[CETESB QUALAR system](https://qualar.cetesb.sp.gov.br/qualar/home.do). This information is often used for air quality model evaluation and air pollution data analysis in Sao Paulo State which is usally perform in R.

## Installation

You can install the released version of qualR by (I hope) :

``` r
devtools:install_github("quishqa/qualR")
```

## Example

### Download one parameter

If you want to download Ozone information from Pinheiros AQS, from January first to January 7th, you can do:

```{r example}
library(qualR)

cetesb_aqs # To check Pinheiros aqs_code
cetesb_param # TO check Ozone pol_code

my_user_name <- "john.doe@mymail.com"
my_password <- "drowssap"
o3_code <- 63
pin_code <- 99
start_date <- "01/01/2020"
end_date <- "07/01/2020"

pin_o3 <- CetesbRetrieve(my_user_name,
                         mymy_password,
                         o3_code,
                         pin_code,
                         start_date,
                         end_date)

```

### Download Criteria Pollutants

```{r example}
pin_pol <- CetesbRetrievePol(my_user_name,
                             my_password,
                             pin_code,
                             start_date,
                             end_date)
```

### Download Meteorological information

```{r example}
pin_met <- CetesbRetrieveMet(my_user_name,
                             my_password,
                             pin_code,
                             start_date,
                             end_date)
```

I hope this will be usefull for your research.
